---
title: "Reported race joint analysis"
author: "Chris McKennan"
date: "4/2/2019"
output: pdf_document
---

Here is the code to estimate the posterior distributions for $(\beta_g^{(0)}, \beta_g^{(7)})$ for each CpG $g =1,\ldots,p=784484$ in the reported race analysis. 

Let $X = A \oplus A \in \mathbb{R}^{2n \times 2}$, where $A \in \mathbb{R}^n$ is each individual's reported race (Black=1, Hispanic=0). Also, let $y_g \in \mathbb{R}^{2n}$ be the methylation and $V_g = \sigma_g^2 I_{2n} + \delta_g^2 B$ be the covariance matrix for the $2n$ individuals at CpG $g$, where $B$ is a partition matrix that blocks samples by individuals. We describe how we estimate $\sigma_g^2$ and $\delta_g^2$ in the Supplement of the manuscript, and define $\hat{V}_g = \hat{\sigma}_g^2 I_{2n} + \hat{\delta}_g^2 B$. Lastly, let $P_M^{\perp}$ be the orthogonal projection matrix that project vectors onto the orthogonal complement of $M$. Our code uses as input the summary statistics
\[
SXX_g = X^TP_{\left[Z\,\hat{C}\right]}^{\perp} \left( P_{\left[Z\,\hat{C}\right]}^{\perp} \hat{V}_g P_{\left[Z\,\hat{C}\right]}^{\perp}\right)^{\dagger} P_{\left[Z\,\hat{C}\right]}^{\perp} X \in \mathbb{R}^{2 \times 2}
\]
and
\[
SXY_g = X^TP_{\left[Z\,\hat{C}\right]}^{\perp} \left( P_{\left[Z\,\hat{C}\right]}^{\perp} \hat{V}_g P_{\left[Z\,\hat{C}\right]}^{\perp}\right)^{\dagger} P_{\left[Z\,\hat{C}\right]}^{\perp} y_g \in \mathbb{R}^2.
\]
The matrix $Z$ contains observed nuisance parameters like the intercept and $\hat{C}$ is an estimate for the latent confounders. We estimated them with a method developed in arXiv:1808.05895v1 (which is currently under a second round revision with JASA: Theory and Methods). We remark that $SXX_g$ and $SXY_g$ are analogous to the univariate summary statistics used in ASH (Stephens, 2016), and that
\[
SXX_g^{-1} SXY_g, \quad SXX_g^{-1}
\]
are the generalized least squares estimate for the bivariate effect due to reported race and its estimated variance.

#Libraries
```{r Libraries}
library(parallel)
library(foreach)
library(mvtnorm)
```

#Source Functions
Make sure to change these paths when before running the code.
```{r SourceFunctions}
source("<path to>/EmpiricalBayes.R")
source("<path to>/PosteriorProbabilities.R")
```

#Load data
Make sure to change these paths when before running the code. SXX is the list of 2 x 2 SXX_g matrices, and SXY is a list of 2-vectors that contains SXY_g.
```{r LoadData}
load("<path to>/SummaryStat_ReportedRace.Rdata")  #Summary statistics necessary to run empirical Bayes and estimate the posterior distribution. These are provided with the manuscript.
```

#Analysis

## Empirical Bayes
out.EB$prior.pi contains the estimated prior
out.EB$Post.probs is a 31 x 784484 matrix, where entry (k,g) is P(Methylation at CpG g was drawn from mixture k | Data). See the Supplement for more details

This code takes a long time to run.
```{r EmpiricalBayes}
var.vec <- c(0.1^2, (0.1 * 3/2)^2, (0.15 * 3/2)^2, (0.2 * 3/2)^2, (0.25 * 3/2)^2)   #A relatively dense grid of effect sizes tau_k^2 (see the Supplement for more detail) obtained from evaluating the effects sizes at birth and age 7 for CpG sites with small RR-qvalues at birth and age 7. A similar technique is done in ASH (Stephens, 2016)
rho.vec <- c(0, 1/3, 2/3, 1)    #The correlation coefficients rho_s (see the Supplement for more detail)
out.EB <- EmBayesPi(SXX=SXX, SXY=SXY, max.iter=5e4, var.vec=var.vec, rho.vec=rho.vec, lambda=100)
```

## Get sign rates
PPS stands for Posterior Probability for Signs, and is used to estimate the conserved and discordant sign rates in the reported race analysis.

By default, the function PostProbs using all but 1 core to determine the posterior sign rates.
```{r SignRates}
PPS <- PostProbs(SXX=SXX, SXY=SXY, Post=out.EB$Post.probs, var.vec=var.vec, rho.vec=rho.vec)
```